A Flux Scale for Southern Hemisphere 21cm EoR Experiments

### 2013-06-18 ms.tex
Incorporating changes from Adrian and Judd. Answering residual notes from Aaron.

Numerous radio telescopes are now exploring the prospects for using
measurements of highly redshifted 21cm emission to inform our understanding of
cosmic reionization in the redshift range $z>6$, corresponding to radio
frequencies below 200 MHz (see reviews in
\citealt{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}).  These
include telescopes aiming to measure the global temperature change of 21cm
emission during the Epoch of Reionization (EoR), such as the Compact
Reionization Experiment (CoRE), the Zero-spacing Interferometer \citep{Raghunathan:2011}  and the Experiment to Detect the Global EoR
Signature (EDGES; \citealt{Bowman:2010p8546}), and interferometers aiming to
measure the power spectrum of 21cm EoR emission, such as the Giant Metre-wave
Radio Telescope (GMRT;
\citealt{Paciga:2011p9470,Paciga:2013p9627})\footnote{\url{http://gmrt.ncra.tifr.res.in/}},
the LOw Frequency ARray (LOFAR;
\citealt{Yatawatta:2013p9699})\footnote{\url{http://www.lofar.org/}}, the
Murchison Widefield Array (MWA;  \citealt{Bowman:2012p9138};
\citealt{Tingay:2013p9022}))\footnote{\url{http://www.mwatelescope.org/}}, and
the Donald C. Backer Precision Array for Probing the Epoch of Reionization
(PAPER; \citealt{Parsons:2010p6757,Pober:2013p9567})\footnote{\url{http://eor.berkeley.edu/}}.

As such, there has been a renewed interest in the spectral and spatial variation of
foreground emission in the 100--200 MHz frequency band that covers
the $z=6$--13 redshift range expected to encompass reionization
\citep{Furlanetto:2006p2267}. % XXX I think this sentence is a repeat from another paper
In particular, the spectral properties of extra-galactic point-sources are important
both because they are valuable calibration references
and because they are strong foreground emitters that must be removed from 21cm EoR measurements.
With the sparse availability of measured foreground properties in the
100--200-MHz frequency band over large areas of the sky \citep{deOliveiraCosta:2008p2242}, continued
foreground characterization is a
vital step en route to any 21cm EoR detection.
At these low
frequencies, the Southern sky is much less well known than the North;
catalog source fluxes at 150MHz are inaccurate at the 20\% level
for $\delta<-20$\arcdeg{} \citep{Slee:1995p7541,Vollmer:2005p6425}.
Both PAPER and the MWA are located in the southern hemisphere at radio-quiet
reserves being prepared for the upcoming Square Kilometer Array, and hence,
the
most extensive surveying work is now being conducted by the EoR experiments
themselves \citep{Jacobs:2011p8438,Williams:2012p8768,Bernardi:2013p9859}.

One significant complication to improving the state of affairs in foreground characterization
is that many 21cm EoR experiments, including LOFAR in the northern hemisphere, and
PAPER and MWA in the southern hemisphere, are designed for drift-scan observations or steered via phased array.
This design decision has largely been driven by the simplicity and cost-effectiveness of
phased and/or correlated dipoles to achieve the aggressive sensitivity requirements
for measuring the 21cm power spectrum of reionization.
%TODO ARP wanted a cite for the above.  I really don't know of a good one. Its it a bit of a broad statement
% XXX how about Parsons 2012a, Beardsley 2012, and jelic 2008 (i think it has sensitivty for lofar)
Adding to the challenge, these telescopes 
cover much wider fields of view ($>10$\arcdeg) and bandwidths
($\sim 100$\% fractional) than traditional dish telescopes.
Because they do not physically point, flux calibration for such arrays relies heavily
on an accurate model of the primary beam response to correct for an
the apparent flux scale that varies across the sky.
This direction dependent gain is currently uncertain to 10\% or higher
\citep{Pober:2012p8800}, and comprises a large fraction of the 20\% difference between current telescopes \citep{Jacobs:2013p9908}.
% XXX difference of what?

@@

The response of the primary beam is of critical importance to EoR measurements.
Differences between the polarization responses cause leakage of polarized
signals into the total intensity measurement possibly corrupting the 
EoR power spectrum \citep{Moore:2013p9621}.  The primary beam 
shape is also critical to measuring and subtracting foregrounds 
\cite{Bernardi:2012p9020,Sullivan:2012p9457,Morales:2012p8790} a process which, to be effective, must be done to better than 1\%
precision for the brightest sources \citet{Liu:2009p4762,Bowman:2009p7816} .
A method for decoupling uncertain fluxes from the uncertain beam has been
described by \citet{Pober:2012p8800}. In simulation the method was able to
achieve 3 to 10\% accuracy in measuring the primary beam, depending on the
number of antennae and other variables. The Pober method emphasized the need
% XXX mention key assumption of 180deg rotational symmetry
for many repeated measurements of each alt-az pointing. Further investigation
is under way to improve and implement this method and would be greatly aided by
the availability of precise flux measurements unaffected by primary beam
uncertainty.

@@

Establishing an accurate spectrum for Pictor A is of particular importance for
PAPER --- a dedicated EoR experiment 
that employs drift-scanning, dual-polarization dipole antennas 
tuned for efficient operation over a 120--170-MHz band.  PAPER is located in the South African Karoo desert
on the Square Kilometer Array
South Africa (SKA-SA) reserve, 100km north of the small town of Carnarvon.
The PAPER array has grown from 16 elements deployed in early 2009 to a
64-element imaging array in 2011 (see Figure \ref{fig:antpos}). 
Since November 2011 it has been arranged in a maximally redundant grid 
configuration to make deep power spectral integrations \citep{Parsons:2012p9028}.
Though highly sensitive as a power spectrum instrument, the maximally redundant array 
has a broad point spread function in the image domain. This severely limits the number 
of sources which may be used for flux calibration. Drift scanning across the sky with a 45\arcdeg{} FWHM
primary beam, there are very few unresolved, bright sources which are far from the galactic plane. 
Pictor A is bright and well enough separated from other emission to dominate the visibilities for a good fraction of
the EoR observing season, making it a desireable source to use for flux calibration.

\section{Approach}
\label{sec:approach}
In this section, we describe our general approach to controlling the impact of beam model errors on measured
spectra derived from 
drift-scan observations 
of Pictor A and a selection of known, bright sources.  Our approach uses a set of ``source tracks'' as
a function of frequency and time.  Each source track is a beam, formed by phasing measured visibilities toward
known source locations as they drift through the primary beam, and summing over antenna pairs, as
described in \S\ref{sec:XXX}.  Since sources fall
at different positions
within the primary beam, it is generally not possible to relate the fluxes of sources at different positions to
one another without an accurate beam model.  To date, the accuracy of source flux measurements in the
southern hemisphere in the 100--200 MHz band have largely been limited by the accuracy of these beam models.

We mitigate this problem by selecting sources within a narrow declination range, so that source tracks represent
nearly identical cuts through the beam response pattern.  Using this fact, the relative amplitudes of sources
can be deduced with minimal reliance on an accurate beam model.  In our analysis, a prior beam model is only
necessary for extrapolating across a narrow declination range (we estimate the errors associated with this 
extrapolation in \S\ref{sec:XXX}), and for approximating optimal inverse-variance weighting when averaging
over source tracks to determine a single spectrum.

This approach addresses several sources of error that were identified in \citet{Jacobs:2013p9908} 
as originating from uncertainty in the primary beam model.
  First, under most imaging schemes, each source flux is measured at just a few 
points in the primary beam. As errors tend to vary across the beam it is often
 difficult to decouple flux uncertainty from beam uncertainty.
 Using the drift scan-beamforming technique, each 
source is measured thousands of times at a variety of different primay beam values to provide a complete sample of the
flux variance due to primary beam variation.  Second, the flux calibration was found to vary over the 
sky due to the dual uncertainties of primary beam response and prior catalog. To reduce our exposure to flux calibration
variation, we limit our observations to sources passing within 5\arcdeg{}  of pointings which we can directly calibrate
to a single reference source.
 A third limitation, identified in \citet{Williams:2012p8768},
was the increase of uncertainty towards the edge of the beam. To minimize our sensitivity to both beam and noise
uncertainty, we weight the source track by an additional factor of the primary beam model.

The remainder of the analysis in this paper relates to calibrating the relative source amplitudes that we
establish to an absolute flux scale as a function of frequency.  Given the shortcomings of 
existing in-band measurements of sources
in the southern hemisphere, we must bootstrap this flux scale from an ensemble of calibrator sources that
exhibit simple power-law spectra.
As described in \S\ref{sec:XXX},
this proceeds by first using a single calibrator to set the relative amplitudes of spectral channels, and
then using several sources, extrapolating across a large set of measurements 
between 70 MHz and 2 GHz, in a Markov-Chain Monte Carlo simulation to establish an absolute flux scale
for the measured spectra.

\subsection{Selection of confirmation sources} % XXX what is meant by "confirmation source" ?  Maybe just "Source Selection"?
As PAPER is a drift scan instrument, each declination describes a distinct path through the primary beam. The flux time seres of a beamform
 provides a detailed sample of the primary beam relative to the peak of the trace, however,
a model is still required to calibrate between declinations. Below we argue that be averaging over long tracks, we minimize our 
susceptibility to localized primary beam error and find that limiting our source selection range to within 
a declination range of 5\arcdeg{} around a calibrator is a good compromise. In \S\ref{fig:gain},
we estimate the error resulting from this limited scope application of the primary beam model.

@@

\subsection{Data Processing Overview}
\label{sec:data_overview}
The spectra reported here are measured by beam-forming ---phasing the visibilities to
the target location and summing over baselines to produce a spectral time series of perceived flux $S_p$,
which is given by the true source flux, $S$, illuminated by the true beam pattern $B$, according to
% XXX would be nice to capture here and below that B, B_M, and S_p are vectors indexed by t.  would
% suggest subscript, except you already have B_M.  Could use B^\prime (and S^\prime) instead...
\begin{equation}
S_p=BS.
\end{equation}
The spectrum is then isolated
from neighboring sources by filtering and averaged in the time domain.  This time-average is 
weighted by a model of PAPER's primary beam response ($B_M$).  
\begin{align}
S_\textrm{est} &= \frac{\sum_t B_M S_p}{\sum_t B^2_M} \nonumber \\ 
&= \frac{\sum_t B_M B}{\sum_t B^2_M} S  \nonumber \\
&= g(\delta) S  \label{eq:g}
\end{align}
% XXX where g and \delta are...?  Should connect this nomeclature to J2331-416 language below (i.e. solving for g)
As the perceived flux is already weighted once by the true beam $B$ this weighted sum is an effective approximation of inverse-variance weighting \citep{Pober:2012p8800}.

The pre-factor in Eq \ref{eq:g} has been integrated in time over the full azimuthal sweep across the sky and therefore depends only on declination. The rate at which this factor
changes with declination defines maximum declination distance we can observe from our calibrator. 
In simulation we multiplied an early holographic beam map \cite{Pober:2012p8800}
% XXX to simulate what?  need to say something like:
% To simulate the effect of beam model errors on ???, we used simulations derived from the holographic
% beam map measured in \citet{Pober:2012p8800} weighted by a beam model derived from electromagnetic
% simulations, and find that ...
by electromagnetic simulations and find that it 
diverges slowly with declination, with a maximum of 10\% over 5\arcdeg{} of declination range, though most are less. 
We estimate this spectral dependent gain at a reference declination of 41\arcdeg{}36\arcmin{}   using (J2331-416) fitting a spectral index model to the
 catalog data points below 2GHz. 
% XXX clarify that this is not the last correction; there is one more flux cal that is derived from MCMC
% over several calibrator sources (Figure XXX).

 Finally we average the spectra from a resolution of 400kHz to 10MHz bins, adding the individual errors in quadrature
 for our uncertainty.

@@

 Measurements are derived from observations using  the east-west dipole arms of
64 PAPER antennas deployed in a minimum-redundancy imaging configuration
(see Figure \ref{fig:antpos})
at the Karoo Radio Observatory site in South Africa.
A100-200MHz band was correlated
with 2048 frequency channels and integrated for 10.7 seconds before
visibilities were stored.  Observations included here on  July 4, 2011, running
from JD2455748.17 to JD2455748.72. and October 17, 2011 running from
JD2455852.2 to JD2455852.6.  In both seasons two antennas were omitted owing to
malfunctioning signal path. This 20 hour long combined dataset provides
observations provides near complete hour angle coverage for the entire 24 hours of
Right Ascension. The July portion of the dataset comes from the same observing
campaign described by \cite{Pober:2013p9567} and \cite{Stefan:2012p9707}. 
%July (\#40,\#55), September (\#7,\#28)

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{plots/auto_tempcorr.png}
\caption{Calibrating the temperature coefficient  between the July and October observing runs.  Here we plot the ratio of auto-correlations (averaged over frequency and antenna) against the temperature difference between the two nights. The very linear middle region has a slope of  -0.055 dB/C while the full, more complex, range is fit by -0.06dB/C (black line).  We compromise by choosing -0.058dB/C   
which was found to significantly improve the agreement (see Figure \ref{fig:auto_compare}).
% XXX When comparing autos from different JDs, the change in position of sun can substantially affect signal level
% I'm worried this is the cause of the dither on either end of Figure 3, unless you've been conservative about
% what data you've included here.
\label{fig:autos_tempcorr}}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\textwidth]{plots/auto_balance.png}
\caption{Gain differences between the July and October nights are dominated by temperature variation. Here we see the affect on 
the average auto correlation before ({\color{orange} orange}) and after (black) application of the temperature dependent gain. \label{fig:auto_compare}}
% XXX why does the gap in data in the top plot not match the bottom?
\end{figure}

Matching of relative gains and phases between antenna was done by fitting a per-antenna complex gain to portions
of data which have easily modeled sky. 
This calibration is only relative between antennae, bandpass and absolute flux calibration
comes after time and frequency averaging in sections \ref{sec:Calibration} and \ref{sec:flux_scale}.   
The antenna delays and amplitudes were found by fitting a point source visibility
model to Centaurus A, Pictor A, and Fornax A.  Each source is imperfectly
modeled by a single point-source, but the solution differences are minimized by
averaging over the three independent solutions. These same calibration
solutions have been successfully applied in \citet{Pober:2013p9567} and \citet{Stefan:2012p9707}.
for power-spectrum analysis of foregrounds and imaging of Centaurus A, respectively.

The October per-antenna gains were found by computing the average ratio of the auto correlations
against the July data. Delays were help constant as in \citet{Jacobs:2011p8438}.

@@

\subsection{Beamforming}

Spectral time-series are computed by beam-forming to the selected sky
locations. A beam is formed by phasing baselines to the desired location and
summing over baselines longer than 20 wavelengths.

@@

The resulting set of spectra were then calibrated to a model of J2331-416
\footnote{Here we set the spectrum to $S_{150}=33$Jy, $\alpha=-0.76$, derived by a linear least squares fit to
catalog measurements below 1GHz \citep{Slee:1995p7541,Kuehr:1981p9628,Large:1981p7798,Burgess:2006p7814}.  Further calibration on top of this somewhat arbitrary calibration model
is explored in detail below. }
 to remove 
the residual bandpass due to the net effect of the primary beam as discussed in section \ref{sec:approach}. 
This source
was selected for its relatively high brightness, spectral smoothness and large quantity
 of available catalog data. Each source track has a slightly different sample profile
 resulting from different amounts of flagged data at each time and channel. To account
 for these differences in the bandpass calibration we build a set of calibration tracks 
 that match the tracks for each source. Calibration of each source then proceeds with
 an optimally matched calibration spectrum.

@@

To model the relationship between different wavelengths we assume a single spectral index
which is the prevailing spectral energy distribution at low frequencies, 
though curvature or other
deviation from a power-law is not-uncommon.

When fitting models to the full catalog, we use the Vollmer et al
catalog which has been optimally cross-matched at the expense of excluding more data points.
Meanwhile, the gain fit used a small sample of
sources with spectra which meet our calibration criteria: more precision data available,
brighter than most sources in the sample and far from any possibly confusing areas of
the sky. 
By limiting ourselves to a small number of sources we are able to go include catalog points by hand,
to avoid making the error of falsely including erroneously matched 
data points but benefiting from some measurements not included.

@@

At the output of the beam forming and bandpass calibration step, the flux scale
 is tied to a model fit of the catalog values of
J2331-416. The accuracy of this fit, and the implied uncertainty in the flux
scale, limits the accuracy of the PAPER measurements..  To refine the flux
scale and estimate our flux scale uncertainty, we bootstrap a single global flux scale 
correction factor using 6
sources selected for their brightness, spectral linearity and data
availability\footnote{Calibration sources:
2250-412,2331-416,1932-464,0103-453,0547-408,0043-424}. To build a more complete
spectral model we go beyond the data found in Vollmer et al to include all spectral
measurements below 2GHz found in the NED database \footnote{\url{ned.ipac.caltech.edu}: Accessed 1 April 2013}.  These additional
 catalog measurements are primarily by Parkes, and Molonglo  \citep{Kuehr:1981p9323} with the best precision coming from the
Wills fluxes at 538 and 634 MHz \citep{Wills:1975p9314}. Where error bars are not
given, we assume uncertainty of 25\%.

@@

Direct measurements of the Pictor A spectrum are key to correctly setting the
flux scale of PAPER, MWA and future EoR experiments like the Hydrogen Epoch of Reionization Array (HERA).
These spectra provide tighter constraints on many of the EoR band fluxes, while
limiting the pernicious affect of primary beam uncertainty.  Future work will use these
fluxes to further refine the primary beam models of these experiments which is crucial to properly reconstructing
both image and power spectrum flux.
Although we have focused on a narrow declination range in this paper, the techniques we describe
may be applied over any declination range.  Future work will also aim to extend this analysis
to other declination ranges, and to tie the flux calibration of these other declination ranges
back to the absolute flux scale we derive here.

@@

% XXX re-check arxiv cites in bib for finishing peer review.  Paciga 2013, Pober 2013, Moore 2013, and Stefan 2013
% all published now.  what is bernardi 2012 submitted?

>>>>

Numerous radio telescopes are now exploring the prospects for using
measurements of highly redshifted 21cm emission to inform our understanding of
cosmic reionization in the redshift range $z>6$, corresponding to radio
frequencies below 200 MHz (see reviews in
\citealt{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}).  These
include telescopes aiming to measure the global temperature change of 21cm
emission during the Epoch of Reionization (EoR), such as the Compact
Reionization Experiment (CoRE), the Zero-spacing Interferometer \citep{Raghunathan:2011}  and the Experiment to Detect the Global EoR
Signature (EDGES; \citealt{Bowman:2010p8546}), and interferometers aiming to
measure the power spectrum of 21cm EoR emission, such as the Giant Metre-wave
Radio Telescope (GMRT;
\citealt{Paciga:2011p9470,Paciga:2013p9943})\footnote{\url{http://gmrt.ncra.tifr.res.in/}},
the LOw Frequency ARray (LOFAR;
\citealt{Yatawatta:2013p9699})\footnote{\url{http://www.lofar.org/}}, the
Murchison Widefield Array (MWA;  \citealt{Bowman:2013p9950};
\citealt{Tingay:2013p9022}))\footnote{\url{http://www.mwatelescope.org/}}, and
the Donald C. Backer Precision Array for Probing the Epoch of Reionization
(PAPER; \citealt{Parsons:2010p6757,Pober:2013p9942})\footnote{\url{http://eor.berkeley.edu/}}.

%As such, there has been a renewed interest in the spectral and spatial variation of
%foreground emission in the 100--200 MHz frequency band that covers
%the $z=6$--13 redshift range expected to encompass reionization
%\citep{Furlanetto:2006p2267}. % XXX I think this sentence is a repeat from another paper
In particular, the spectral properties of extra-galactic point-sources are important
both because they are valuable calibration references
and because they are strong foreground emitters that must be removed from 21cm EoR measurements.
With the sparse availability of measured foreground properties in the
100--200-MHz frequency band over large areas of the sky \citep{deOliveiraCosta:2008p2242}, continued
foreground characterization is a
vital step en route to any 21cm EoR detection.
At these low
frequencies, the Southern sky is much less well known than the North;
catalog source fluxes at 150MHz are inaccurate at the 20\% level
for $\delta<-20$\arcdeg{} \citep{Slee:1995p7541,Vollmer:2005p6425}.
Both PAPER and the MWA are located in the southern hemisphere at radio-quiet
reserves being prepared for the upcoming Square Kilometer Array, and hence,
the
most extensive surveying work is now being conducted by the EoR experiments
themselves \citep{Jacobs:2011p8438,Williams:2012p8768,Bernardi:2013p9859}.

One significant complication to improving the state of affairs in foreground characterization
is that many 21cm EoR experiments, including LOFAR in the northern hemisphere, and
PAPER and MWA in the southern hemisphere, are designed for drift-scan observations or steered via phased array.
This design decision has largely been driven by the simplicity and cost-effectiveness of
phased and/or correlated dipoles to achieve the aggressive sensitivity requirements
for measuring the 21cm power spectrum of reionization \citep{Parsons:2012p9028,Beardsley:2012p9136,Jelic:2008p2130}.
Adding to the challenge, these telescopes 
cover much wider fields of view ($>10$\arcdeg) and bandwidths
($\sim 100$\% fractional) than traditional dish telescopes.
Because they do not physically point, flux calibration for such arrays relies heavily
on an accurate model of the primary beam response to correct for an
the apparent flux scale that varies across the sky.
This direction dependent gain is currently uncertain to 10\% or higher
\citep{Pober:2012p8800}, and comprises a large fraction of the 20\% flux uncertainty between current telescopes \citep{Jacobs:2013p9908}.

@@

The response of the primary beam is of critical importance to EoR measurements.
Differences between the polarization responses cause leakage of polarized
signals into the total intensity measurement possibly corrupting the 
EoR power spectrum \citep{Moore:2013p9941}.  The primary beam 
shape is also critical to measuring and subtracting foregrounds 
\cite{Bernardi:2013p9859,Sullivan:2012p9457,Morales:2012p8790} a process which, to be effective, must be done to better than 1\%
precision for the brightest sources \citet{Liu:2009p4762,Bowman:2009p7816} .
A method for decoupling uncertain fluxes from the uncertain beam has been
described by \citet{Pober:2012p8800}. In simulation the method was able to
achieve 3 to 10\% accuracy in measuring the primary beam, depending on the
number of antennae and other variables. In particular, it emphasized the need
for many repeated measurements of each alt-az pointing, which in that
case were found by assuming 180\arcdeg{} symmetry. Further investigation
is under way to improve and implement this method and would be greatly aided by
the availability of precise flux measurements unaffected by primary beam
uncertainty.

@@

Establishing an accurate spectrum for Pictor A is of particular importance for
PAPER --- a dedicated EoR experiment 
that employs drift-scanning, dual-polarization dipole antennas 
tuned for efficient operation over a 120--170-MHz band.  PAPER is located in the South African Karoo desert
on the Square Kilometer Array
South Africa (SKA-SA) reserve, 100km north of the small town of Carnarvon.
The PAPER array has grown from 16 elements deployed in early 2009 to a
64-element imaging array in 2011 (see Figure \ref{fig:antpos}). 
Since November 2011 it has been arranged in a maximally redundant grid 
configuration to make deep power spectral integrations \citep{Parsons:2012p9028}.
Though highly sensitive as a power spectrum instrument, the maximally redundant array 
has a broad point spread function in the image domain. This severely limits the number 
of sources which may be used for flux calibration. Drift scanning across the sky with a 45\arcdeg{} FWHM
primary beam, there are very few unresolved, bright sources which are far from the galactic plane. 
Pictor A is bright and well enough separated from other emission to dominate the visibilities for a good fraction of
the EoR observing season, making it a desirable source to use for flux calibration.

\section{Approach}
\label{sec:approach}
In this section, we describe our general approach to controlling the impact of beam model errors on measured
spectra derived from 
drift-scan observations 
of Pictor A and a selection of known, bright sources.  Our approach uses a set of ``source tracks'' as
a function of frequency and time.  Each source track is a beam, formed by phasing measured visibilities toward
known source locations as they drift through the primary beam, and summing over antenna pairs, as
described in \S\ref{sec:beamforming}.  Since sources fall
at different positions
within the primary beam, it is generally not possible to relate the fluxes of sources at different positions to
one another without an accurate beam model.  To date, the accuracy of source flux measurements in the
southern hemisphere in the 100--200 MHz band have largely been limited by the accuracy of these beam models.

We mitigate this problem by selecting sources within a narrow declination range, so that source tracks represent
nearly identical cuts through the beam response pattern.  Using this fact, the relative amplitudes of sources
can be deduced with minimal reliance on an accurate beam model.  In our analysis, a prior beam model is only
necessary for extrapolating across a narrow declination range (we estimate the errors associated with this 
extrapolation in \S\ref{sec:flux_scale}), and for approximating optimal inverse-variance weighting when averaging
over source tracks to determine a single spectrum.

This approach addresses several sources of error that were identified in \citet{Jacobs:2013p9908} 
as originating from uncertainty in the primary beam model.
  First, under most imaging schemes, each source flux is measured at just a few 
points in the primary beam. As errors tend to vary across the beam it is often
 difficult to decouple flux uncertainty from beam uncertainty.
 Using the drift scan-beamforming technique, each 
source is measured thousands of times at a variety of different primary beam values to provide a complete sample of the
flux variance due to primary beam variation.  Second, the flux calibration was found to vary over the 
sky due to the dual uncertainties of primary beam response and prior catalog. To reduce our exposure to flux calibration
variation, we limit our observations to sources passing within 5\arcdeg{}  of pointings which we can directly calibrate
to a single reference source.
 A third limitation, identified in \citet{Williams:2012p8768},
was the increase of uncertainty towards the edge of the beam. To minimize our sensitivity to both beam and noise
uncertainty, we weight the source track by an additional factor of the primary beam model.

The remainder of the analysis in this paper relates to calibrating the relative source amplitudes that we
establish to an absolute flux scale as a function of frequency.  Given the shortcomings of 
existing in-band measurements of sources
in the southern hemisphere, we must bootstrap this flux scale from an ensemble of calibrator sources that
exhibit simple power-law spectra.
As described in \S\ref{sec:flux_scale},
this proceeds by first using a single calibrator to set the relative amplitudes of spectral channels, and
then using several sources, extrapolating across a large set of measurements 
between 70 MHz and 2 GHz, in a Markov-Chain Monte Carlo simulation to establish an absolute flux scale
for the measured spectra.

\subsection{Source Selection} 
As PAPER is a drift scan instrument, each declination describes a distinct path through the primary beam. The flux time series of a beamform
 provides a detailed sample of the primary beam relative to the peak of the trace, however,
a model is still required to calibrate between declinations. Below we argue that by averaging over long tracks, we minimize our 
susceptibility to localized primary beam error and find that limiting our source selection range to within 
a declination range of 5\arcdeg{} around a calibrator is a good compromise. In \S\ref{fig:gain},
we estimate the error resulting from this limited scope application of the primary beam model.

@@

\subsection{Data Processing Overview}
\label{sec:data_overview}
The spectra reported here are measured by beam-forming ---phasing the visibilities to
the target location and summing over baselines to produce a spectral time series of perceived flux $S_p$,
which is given by the true source flux, $S$, illuminated by the true beam pattern $B$, according to
\begin{equation}
S_p=B(t)S.
\end{equation}
The spectrum is then isolated
from neighboring sources by filtering and averaged in the time domain.  This time-average is 
weighted by a model of PAPER's primary beam response ($B_M$). As the perceived flux is already weighted once by the true beam $B$ this weighted sum is an effective approximation of inverse-variance weighting \citep{Pober:2012p8800}. 
\begin{align}
S_\textrm{est} &= \frac{\sum_t B(t)_M S_p}{\sum_t B(t)^2_M} \nonumber \\ 
&= \frac{\sum_t B(t)_M B(t)}{\sum_t B(t)^2_M} S  \nonumber \\
&\equiv g(\delta) S  \label{eq:g}
\end{align}
The net time integrated, weighted beam response $g$ is purely a function of declination $\delta$.The rate at which this factor
changes with declination defines maximum declination distance we can observe from our calibrator. 
To simulate this affect we we multiplied a measurement of the primary beam \citep{Pober:2012p8800}
by electromagnetic simulations \citep{Parsons:2010p6757} and find that $g$ 
diverges slowly with declination, with a maximum of 10\% over 5\arcdeg{} of declination range, though most are less. Thus, to first order,
calibration to a single declination will remove most of the variation with declination.
We estimate this spectral dependent $g$ at a reference declination of 41\arcdeg{}36\arcmin{}  using (J2331-416) fitting a spectral index model to the
 catalog data points below 2GHz. This levels the bandpass due to spectral dependence  of the primary bean 
 and puts all the sources on approximately the same flux scale.  
 Finally we average the spectra from a resolution of 400kHz to 10MHz bins, adding the individual errors in quadrature
 for our uncertainty.

@@

 Measurements are derived from observations using two nights on the east-west dipole arms of
64 PAPER antennas deployed in a minimum-redundancy imaging configuration
(see Figure \ref{fig:antpos})
at the Karoo Radio Observatory site in South Africa.
A100-200MHz band was correlated
with 2048 frequency channels and integrated for 10.7 seconds before
visibilities were stored.  Observations included here on  July 4, 2011, running
from JD2455748.17 to JD2455748.72. and October 17, 2011 running from
JD2455852.2 to JD2455852.6.  In both seasons two antennas were omitted owing to
malfunctioning signal path. This 20 hour long combined dataset provides
observations provides near complete hour angle coverage for the entire 24 hours of
Right Ascension. The July portion of the dataset comes from the same observing
campaign described by \cite{Pober:2013p9942} and \cite{Stefan:2013p9926}. 
%July (\#40,\#55), September (\#7,\#28)

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{plots/auto_tempcorr.png}
\caption{Calibrating the temperature coefficient  between the July and October nights.  Here we plot the ratio of auto-correlations (averaged over frequency and antenna) against the temperature difference between the two nights. The very linear middle region has a slope of  -0.055 dB/C while the full, more complex, range is fit by -0.06dB/C (black line).  We compromise by choosing -0.058dB/C   
which was found to significantly improve the agreement (see Figure \ref{fig:auto_compare}).
% XXX When comparing autos from different JDs, the change in position of sun can substantially affect signal level
% I'm worried this is the cause of the dither on either end of Figure 3, unless you've been conservative about
% what data you've included here.
% There is no sun data.  Changed "observing runs" to "nights". Is that what you meant? --DCJ
\label{fig:autos_tempcorr}}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\textwidth]{plots/auto_balance.png}
\caption{Gain differences between the July and October nights are dominated by temperature variation. Here we see the effect on 
the average auto correlation before ({\color{orange} orange}) and after (black) application of the temperature dependent gain. \label{fig:auto_compare}}.   Note that the span of the data in the bottom fractional difference plot is limited to points occurring in both nights and does not cover the full range seen in the top.
\end{figure}

Matching of relative gains and phases between antenna was done by fitting a per-antenna complex gain to portions
of data which have easily modeled sky. 
This calibration is only relative between antennae, bandpass and absolute flux calibration
comes after time and frequency averaging in sections \ref{sec:Calibration} and \ref{sec:flux_scale}.   
The antenna delays and amplitudes were found by fitting a point source visibility
model to Centaurus A, Pictor A, and Fornax A.  Each source is imperfectly
modeled by a single point-source, but the solution differences are minimized by
averaging over the three independent solutions. These same calibration
solutions have been successfully applied in \citet{Pober:2013p9942} and \citet{Stefan:2013p9926}.
for power-spectrum analysis of foregrounds and imaging of Centaurus A, respectively.

The October per-antenna gains were found by computing the average ratio of the auto correlations
against the July data. Delays were held constant as in \citet{Jacobs:2011p8438}.

@@

\subsection{Beamforming}
\label{sec:beamforming}
Spectral time-series are computed by beam-forming to the selected sky
locations. A beam is formed by phasing baselines to the desired location and
summing over baselines longer than 20 wavelengths.

@@

The resulting set of spectra were then calibrated to a model of J2331-416 
\footnote{Here we set the spectrum to $S_{150}=33$Jy, $\alpha=-0.76$, derived by a linear least squares fit to
catalog measurements below 1GHz \citep{Slee:1995p7541,Kuehr:1981p9628,Large:1981p7798,Burgess:2006p7814}.  Further calibration on top of this somewhat arbitrary calibration model
is explored in detail below. }
 to remove 
the residual bandpass $g$ from equation \ref{eq:g} caused by the net effect of the primary beam as discussed in section \ref{sec:approach}. 
This source
was selected for its relatively high brightness, spectral smoothness and large quantity
 of available catalog data. Each source track has a slightly different sample profile
 resulting from different amounts of flagged data at each time and channel. To account
 for these differences in the bandpass calibration we build a set of calibration tracks 
 that match the tracks for each source. Calibration of each source then proceeds with
 an optimally matched calibration spectrum.

@@

To model the relationship between different wavelengths we assume a single spectral index
which is the prevailing spectral energy distribution at low frequencies, 
though curvature or other
deviation from a power-law is not uncommon.

When fitting models to the full catalog, we use the Vollmer et al
catalog which has been optimally cross-matched at the expense of excluding more data points.
Meanwhile, the gain fit used a small sample of
sources with spectra which meet our calibration criteria: more precision data available,
brighter than most sources in the sample and far from any possibly confusing areas of
the sky. 
By limiting ourselves to a small number of sources we are able to go include catalog points by hand,
to avoid making the error of falsely including erroneously matched 
data points.

@@

At the output of the beam forming and bandpass calibration step, the flux scale
 is tied to a model fit of the catalog values of
J2331-416. The accuracy of this fit, and the implied uncertainty in the flux
scale, limits the accuracy of the PAPER measurements.  To refine the flux
scale and estimate our flux scale uncertainty, we bootstrap a single global flux scale 
correction factor using 6
sources selected for their brightness, spectral linearity and data
availability\footnote{Calibration sources:
2250-412,2331-416,1932-464,0103-453,0547-408,0043-424}. To build a more complete
spectral model we go beyond the data found in Vollmer et al to include all spectral
measurements below 2GHz found in the NED database \footnote{\url{ned.ipac.caltech.edu}: Accessed 1 April 2013}.  These additional
 catalog measurements are primarily by Parkes, and Molonglo  \citep{Kuehr:1981p9323} with the best precision coming from the
Wills fluxes at 538 and 634 MHz \citep{Wills:1975p9314}. Where error bars are not
given, we assume uncertainty of 25\%.

@@

Direct measurements of the Pictor A spectrum are key to correctly setting the
flux scale of PAPER, MWA and future EoR experiments like the Hydrogen Epoch of Reionization Array (HERA).
These spectra provide tighter constraints on many of the EoR band fluxes, while
limiting the pernicious effect of primary beam uncertainty.  Future work will use these
fluxes to further refine the primary beam models of these experiments which is crucial to properly reconstructing
both image and power spectrum flux.
Although we have focused on a narrow declination range in this paper, the techniques we describe
may be applied over any declination range.  Future work will also aim to extend this analysis
to other declination ranges, and to tie the flux calibration of these other declination ranges
back to the absolute flux scale we derive here.

@@

% all published now.  what is bernardi 2012 submitted?
